Mechanization of Separation in Generic Extensions

### 2019-01-09 LICS/hacking.tex
guión en non-empty
varias observaciones de Manu

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Hacking of \isatt{ZF-Constructible}}
\label{sec:hack-constructible}
In \cite{paulson_2003}, Paulson presented his formalization of the
relative consistency of the Axiom of Choice. This development is
included inside the Isabelle distribution under the session 
\isatt{ZF-Constructible}. The main technical devices, invented by
G\"odel for this purpose, are \emph{relativization} and
\emph{absoluteness}. In a nutshell, to relativize a formula $\phi$ to
a class $C$, it is enough to restrict its quantifiers to $C$. The
example of \isatt{upair\_ax} in
Section~\ref{sec:axioms-models-set-theory}, the relativized version of
the Pairing Axiom, is extracted from \texttt{Relative}, one of the
core theories of \isatt{ZF-Constructible}. On the other hand, $\phi$
is \emph{absolute} for $C$ if it is equivalent to its relativization,
meaning that the statement made $\phi$ coincides with what $C$
``believes'' $\phi$ is saying. Paulson shows that under certain
hypotheses  on a class $M$ (condensed in the locale \isatt{M\_trivial}), a plethora of
absoluteness and closure results can be proved about $M$.

@@

The most notable changes, located in the theories \texttt{Relative}
and \isatt{WF\_absolute}, are
the following:
\begin{enumerate}
\item\label{item:1} The locale \isatt{M\_trivial}
  does not assume that the underlying class $M$ satisfies the relative
  Axiom of replacement.  As a consequence, the lemma
  \isatt{strong\_replacementI} is no longer valid and was commented
  out.
\item\label{item:2}  Originally the Powerset Axiom was assumed by the
  locale \isatt{M\_trivial},   we moved this requirement to \isatt{M\_basic}. 
\item\label{item:3} We replaced the need that the set of natural
  numbers is in $M$ by the   milder hypothesis that $M(0)$. Actually,
  most results should follow 
  by only assuming that $M$ is nonempty.
\item We moved the requirement $M(\mathtt{nat})$ to the locale
  \isatt{M\_trancl}, where it is needed for the first time. Some results,
  for instance \isatt{rtran\_closure\_mem\_iff} and 
  \isatt{iterates\_imp\_wfrec\_replacement} had to be moved inside that
  locale. 
\end{enumerate}
Because of these changes, some theory files from the
\isatt{ZF-Constructible} session have been included among ours.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%          
\section{Hacking of \isatt{ZF-Constructible}}
\label{sec:hack-constructible}
In \cite{paulson_2003}, Paulson presented his formalization of the
relative consistency of the Axiom of Choice. This development is
included inside the Isabelle distribution under the session 
\isatt{ZF-Constructible}. The main technical devices, invented by
G\"odel for this purpose, are \emph{relativization} and
\emph{absoluteness}. In a nutshell, to relativize a formula $\phi$ to
a class $C$, it is enough to restrict its quantifiers to $C$. The
example of \isatt{upair\_ax} in
Section~\ref{sec:axioms-models-set-theory}, the relativized version of
the Pairing Axiom, is extracted from \texttt{Relative}, one of the
core theories of \isatt{ZF-Constructible}. On the other hand, $\phi$
is \emph{absolute} for $C$ if it is equivalent to its relativization,
meaning that the statement made by $\phi$ coincides with what $C$
``believes'' $\phi$ is saying. Paulson shows that under certain
hypotheses  on a class $M$ (condensed in the locale \isatt{M\_trivial}), a plethora of
absoluteness and closure results can be proved about $M$.

@@

The most notable changes, located in the theories \texttt{Relative}
and \isatt{WF\_absolute}, are
the following:
\begin{enumerate}
\item\label{item:1} The locale \isatt{M\_trivial}
  does not assume that the underlying class $M$ satisfies the relative
  Axiom of replacement.  As a consequence, the lemma
  \isatt{strong\_replacementI} is no longer valid and was commented
  out.
\item\label{item:2}  Originally the Powerset Axiom was assumed by the
  locale \isatt{M\_trivial},   we moved this requirement to \isatt{M\_basic}. 
\item\label{item:3} We replaced the need that the set of natural
  numbers is in $M$ by the   milder hypothesis that $M(0)$. Actually,
  most results should follow 
  by only assuming that $M$ is non-empty.
\item We moved the requirement $M(\mathtt{nat})$ to the locale
  \isatt{M\_trancl}, where it is needed for the first time. Some results,
  for instance \isatt{rtran\_closure\_mem\_iff} and 
  \isatt{iterates\_imp\_wfrec\_replacement} had to be moved inside that
  locale. 
\end{enumerate}
Because of these changes, some theory files from the
\isatt{ZF-Constructible} session have been included among ours.
